Space Matters: extending sensitivity analysis to initial spatial conditions in geosimulation models

### 2019-09-02 paper.tex
submit_v2

%\comment[FL]{le titre de l'article ce serait presque "geometry" matters? non?} bof ca parle moins

@@

%%%%%%%
%% Highlights
%
% - We investigate the sensitivity of geosimulation models to initial spatial conditions.
% - We generate stylized population density grids.
% - We assess the sensitivity of two classical agent-based models, Schelling's model and Sugarscape's model.
% - Strong quantitative and qualitative effects of spatial initial conditions are shown for these models.

@@

Sugarscape is a model of resource extraction which simulates the unequal distribution of wealth within a heterogenous population \cite{EpsteinAxtell1996}. Although it "is designed to study the interaction of many plausible social mechanisms" \cite[p.125]{Axtelletal1996}, we refer in this paper to the first (and simplest) version of the model, where "processes allow its agents to look for, move to, and eat a resource ("sugar") which grows on its $[...]$ array of cells". Agents of different vision scopes and different metabolisms harvest a self-regenerating resource available heterogeneously in the initial landscape, they settle and collect this resource, which leads some of them to survive and others to perish. The main parameters of this model are the number of agents, their minimal and maximal resource levels. In an urban environment, Sugarscape can be used to model how the spatial distribution of any type of goods or services can influence the spread of wealth among inhabitants. Following \cite{batty2005agents}, it can be considered as a metaphor of an urban system. We extend the implementation with agents wealth distribution of \cite{li2009netlogo}. The outcome of the model is measured by a Gini index of inequality for resource distribution. We are interested in testing the impact of the spatial distribution of the resource, using the generated grids.

>>>>

@@

@@

Sugarscape is a model of resource extraction which simulates the unequal distribution of wealth within a heterogenous population \citep{EpsteinAxtell1996}. Although it "is designed to study the interaction of many plausible social mechanisms" \cite[p.125]{Axtelletal1996}, we refer in this paper to the first (and simplest) version of the model, where "processes allow its agents to look for, move to, and eat a resource ("sugar") which grows on its $[...]$ array of cells". Agents of different vision scopes and different metabolisms harvest a self-regenerating resource available heterogeneously in the initial landscape, they settle and collect this resource, which leads some of them to survive and others to perish. The main parameters of this model are the number of agents, their minimal and maximal resource levels. In an urban environment, Sugarscape can be used to model how the spatial distribution of any type of goods or services can influence the spread of wealth among inhabitants. Following \cite{batty2005agents}, it can be considered as a metaphor of an urban system. We extend the implementation with agents wealth distribution of \cite{li2009netlogo}. The outcome of the model is measured by a Gini index of inequality for resource distribution. We are interested in testing the impact of the spatial distribution of the resource, using the generated grids.
